\vskip-0.6in Temporary Trade Barriers: When Will They End?

### 2017-11-10 draft.tex
rework extended abstract into updated intro

This question can be addressed in the context of renewals of trade remedies under the WTO agreements. Normally, temporary measures such as anti-dumping and counter-vailing duties are imposed with a five-year sunset provision that is subject to renewal by the U.S. International Trade Commission. Conditional on a temporary trade barrier being granted and applied, one would like to know what determines whether a renewal order is granted so that the measure continues in force.

In many countries, the crucial actors in the imposition of temporary trade barriers are advocates for the industry that would be protected and an administrative body that has been granted authority to make decisions about whether a given temporary trade barrier is WTO-legal. In the United States, this body is the International Trade Commission (ITC). For ease of exposition and because I would ultimately like to test this theory using U.S. data, I will refer to the actors in this model as industries and the ITC.

In this model, there are two phases of interaction between any given industry seeking protection and the ITC. In both cases, I assume that all actors take most-favored-nation (MFN) tariffs as given.

In the first phase, when no temporary trade barrier is in place, an industry exerts effort to convince the ITC to enact a temporary trade barrier for its product. The ITC then must decide whether to maintain the MFN tariff or impose the temporary trade barrier. Importantly, we assume that the decisions of the ITC are not deterministic, with the amount of uncertainty varying by industry.

Given that the industry receives temporary protection in the first phase of interaction, there will be another phase of interaction when the trade barrier expires, here modeled as five years later to match the typical five years sunset provision under the WTO. At this point, the interaction is similar to that during the first phase, with the industry first exerting effort and then the ITC deciding whether to extend the temporary trade barrier or to revert to the MFN tariff. The essential strategic difference is that at this point the level of the temporary trade barrier is taken as exogenously set at the level determined in the first round of interaction.

An industry's decision about how much, if any, effort to exert depends on a number of factors. Among them are the gap between the applied tariff it faces and the protection it would receive under a temporary trade barrier, the cost of seeking the temporary trade barrier, and, crucially, the probability that its request will be granted. That is, the industry's incentives to seek protection, and the intensity with which it does so, depend on how much uncertainty it faces in the ITC's decision-making process.

Thus we turn to the question of how the ITC decides whether to grant a temporary trade barrier. The imposition of a temporary trade barrier, or continuation of one already in place, is not in general costless. We might assume then that the ITC only grants renewals when it finds the benefits outweigh the costs. If the ITC's rulings are uncertain from the point of view of industry, it must be that the industry cannot perfectly predict the outcome of any given proceeding. That is, the industry does not know with precision how the ITC weighs the costs and benefits of granting protection.

There are several possible sources of this uncertainty. The industrial group may not be able to predict directly the quality of evidence the ITC requires in order to be willing to provide protection through a temporary trade barrier. Indirectly, this may reflect the ITC's preferences about the costs of disputes and retaliation by trade partners. Industry may also not be fully informed about how the ITC weighs the potential harms to the industry versus those to up- and down-stream industries, consumers and trading relations. This may be derived from factors such as the political influence of the industry, how central it is to the economy, and whether there is active lobbying by producers downstream of the product in question.

\newpage
\subsection{Outline}
\begin{enumerate}
  \item What are the determinants of renewals?
\end{enumerate}

\vskip.2in
\begin{itemize}
	\item Temporary trade barriers have potential to be renewed
		\begin{itemize}
			\item For Anti-dumping (AD) duties, initial five year term, renewal for five years
			\item For Safeguards, four then four; much rarer than AD
		\end{itemize}
	\item Authorized by WTO
	\item Very little in the agreements to guide the process
		\begin{itemize}
			\item WTO litigation gives gov'ts a LOT of latitude
			\item Renewal decision determined by complex process including DOC, ITC, pressure via other political bodies.
		\end{itemize}
	\item In U.S., for AD:
		\begin{itemize}
			\item DOC initiates, determines whether dumping would continue
						\begin{itemize}
			\item Can renew AD duties
			\item Susceptible to influence of lobbying, perhaps both direct and indirect
			\item Modeled in reduced form
		\end{itemize}

>>>>

This question can be addressed in the context of renewals of trade remedies under the WTO agreements. Normally, temporary measures such as anti-dumping and counter-vailing duties are imposed with a sunset provision that is subject to renewal. Conditional on a temporary trade barrier being granted and applied, one would like to know what determines whether a renewal order is granted so that the measure continues in force. In fact, this very particular institutional environment allows us to examine questions about the determinants of protection that are usually much more difficult to answer by holding fixed some of the key elements.

In many countries, the crucial actors in the imposition of temporary trade barriers are advocates for the industry that would be protected and some constellation of administrative bodies that have been granted authority to make decisions about whether a given temporary trade barrier is WTO-legal. For ease of exposition and because I would ultimately like to test this theory using U.S. data, I will refer to the actors in this model as industries and the government, and I will focus on antidumping duties because antidumping duties are both imposed and renewed most frequently and thus provided the largest sample size.

In this model, there are two phases of interaction between any given industry seeking protection and the government. In both cases, I assume that all actors take most-favored-nation (MFN) tariffs as given.

In the first phase, when no temporary trade barrier is in place, an industry exerts effort to convince the government to enact a temporary trade barrier for its product. The government must then decide whether to maintain the MFN tariff or impose the temporary trade barrier. Importantly, we assume that the decisions of the government are not deterministic, with the amount of uncertainty varying by industry.

Given that the industry receives temporary protection in the first phase of interaction, there will be another phase of interaction when the trade barrier expires, here modeled as five years later to match the typical five years sunset provision for antidumping duties under the WTO. At this point, the interaction is similar to that during the first phase, with the industry first exerting effort and then the ITC deciding whether to extend the temporary trade barrier or to revert to the MFN tariff. The essential strategic difference is that at this point the level of the temporary trade barrier is taken as exogenously set at the level determined in the first round of interaction. It is the second phase that is modeled in the current paper, with the addition of the first phase acknowledged as a very important and imminent extension.

An industry's decision about how much, if any, effort to exert depends on a number of factors. Among them are the gap between the applied tariff it faces and the protection it would receive under a temporary trade barrier, the cost of seeking the temporary trade barrier, and, crucially, the probability that its request will be granted. That is, the industry's incentives to seek protection, and the intensity with which it does so, depend on how much uncertainty it faces in the government's decision-making process.

Thus we turn to the question of how the government decides whether to grant a temporary trade barrier. The imposition of a temporary trade barrier, or continuation of one already in place, is not in general costless. We might assume then that the government only grants renewals when it finds the benefits outweigh the costs. If the government's rulings are uncertain from the point of view of industry, it must be that the industry cannot perfectly predict the outcome of any given proceeding. That is, the industry does not know with precision how the government weighs the costs and benefits of granting protection.

There are several possible sources of this uncertainty. The industrial group may not be able to predict directly the quality of evidence the government requires in order to be willing to provide protection through a temporary trade barrier. Indirectly, this may reflect the government's preferences about the costs of disputes and retaliation by trade partners. Industry may also not be fully informed about how the government weighs the potential harms to the industry versus those to up- and down-stream industries, consumers and trading relations. This may be derived from factors such as the political influence of the industry, how central it is to the economy, and whether there is active lobbying by producers downstream of the product in question.

\subsection{Outline}

\vskip.2in
\begin{itemize}
	\item Authorized by WTO
	\item Very little in the agreements to guide the process
		\begin{itemize}
			\item WTO litigation gives gov'ts a LOT of latitude
			\item Renewal decision determined by complex process including DOC, ITC, pressure via other political bodies.
		\end{itemize}
	\item In U.S., for AD:
		\begin{itemize}
			\item DOC initiates, determines whether dumping would continue
						\begin{itemize}
			\item Can renew AD duties
			\item Susceptible to influence of lobbying, perhaps both direct and indirect
			\item Modeled in reduced form
		\end{itemize}
